Poll-Based Forecasting of the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd
typo fixed
typo fixed

---
title: "Poll-Based Forecasting of the 2024 U.S. Presidential Election"
subtitle: "Valid Multiple Linear Regression Explains Why Kamala Harris Will Win"
author: 
  - Haowei Fan
  - Fangning Zhang
  - Shaotong Li
thanks: "Code and data are available at: [https://github.com/HaoweiFan0912/US_Election-Forecast/tree/main]."
date: today
date-format: long
abstract: "The United States plays a major role in the global technological and economic landscape, with its political leadership having a significant influence on international technology cooperation and economic relations. This project uses multiple linear regression to predict the outcome of the 2024 U.S. presidential election by extracting and analyzing polling data from polls that nominated the same candidates, providing direction to help stakeholders prepare for potential shifts in U.S. policies. The project predicts that Kamala Harris will secure 47.6% of the vote over Donald Trump's 43.5% and presents a straightforward yet effective regression model that does not rely on time-series analysis or external events."
format: pdf
toc-depth: 2    
number-sections: true
toc: true           
toc-title: "Table of Contents" 
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false
#### Install packages ####
required_packages <- c("tidyverse","knitr","dplyr","arrow","patchwork","car","kableExtra","gridExtra","moments")
for (pkg in required_packages) {
  if (!requireNamespace(pkg,quietly = TRUE)) {
    install.packages(pkg)
  }
}

@@

First, we verified the assumptions mentioned in the @Model_Overview. The @tbl-vohfm shows the General Variance Inflation Factor (GVIF) for both models, which indicates that all predictors have a GVIF less than 1.3, suggesting no significant multicollinearity.

@@

The following @fig-mdp presents the diagnostic plots for the Harris and Trump's models. The plots on the left shows the relationship between residuals and predicted values, and since no obvious pattern is observed, our models satisfies the Homoscedasticity assumption. The right-hand plot are a Q-Q plot, and the points align along the diagonal, indicating that residuals are normally distributed and satisfy the Multivariate Normality condition.

@@

# Result {#sec-result}
## featured values used in prediction

@@

### Create Plots for Numeric Variables ###
# Create a list to store plots
plots <- list()
# Loop through selected numeric variables and plot their distributions
for (variable in set3_vars) {
  plots[[variable]] <- ggplot(raw_data, aes_string(x = variable)) +
    geom_bar(fill = "#69b3a2", color = "black", alpha = 0.8) +
    labs(title = paste("Distribution of", variable), x = variable, y = "Frequency") +
    theme_minimal(base_size = 30) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 10)
    )
}
### Combine and Print Plots ###
# Combine all individual plots into one using patchwork with 2 columns and 4 rows
combined_plot <- wrap_plots(plots, ncol = 2, nrow = 2) +
  plot_annotation(
    theme = theme(
      plot.title = element_text(hjust = 0.5, size = 14)
    )
  )
# Print the combined plot
print(combined_plot)
```
# References

>>>>

---
title: "Poll-Based Forecasting of the 2024 U.S. Presidential Election"
subtitle: "Valid Multiple Linear Regression Explains Why Kamala Harris Will Win"
author: 
  - Haowei Fan
  - Fangning Zhang
  - Shaotong Li
thanks: "Code and data are available at: https://github.com/HaoweiFan0912/US_Election-Forecast/tree/main."
date: today
date-format: long
abstract: "The United States plays a major role in the global technological and economic landscape, with its political leadership having a significant influence on international technology cooperation and economic relations. This project uses multiple linear regression to predict the outcome of the 2024 U.S. presidential election by extracting and analyzing polling data from polls that nominated the same candidates, providing direction to help stakeholders prepare for potential shifts in U.S. policies. The project predicts that Kamala Harris will secure 47.6% of the vote over Donald Trump's 43.5% and presents a straightforward yet effective regression model that does not rely on time-series analysis or external events."
format: pdf
toc-depth: 2    
number-sections: true
toc: true           
toc-title: "Table of Contents" 
bibliography: references.bib
---
```{r}
#| include: false
#| warning: false
#| message: false
#| echo: false
#### Install packages ####
required_packages <- c("tidyverse","knitr","dplyr","arrow","patchwork","car","kableExtra","gridExtra","moments")
for (pkg in required_packages) {
  if (!requireNamespace(pkg,quietly = TRUE)) {
    install.packages(pkg)
  }
}

@@

First, we verified the assumptions mentioned in the @sec-Model_Overview. The @tbl-vohfm shows the General Variance Inflation Factor (GVIF) for both models, which indicates that all predictors have a GVIF less than 1.3, suggesting no significant multicollinearity.

@@

The following @fig-mdp presents the diagnostic plots for the Harris and Trump's models. The plots on the left shows the relationship between residuals and predicted values, and since no obvious pattern is observed, our models satisfies the Homoscedasticity assumption. The right-hand plot are Q-Q plots, and the points align along the diagonal, indicating that residuals are normally distributed and satisfy the Multivariate Normality condition.

@@

# Result {#sec-result}
## Featured values used in prediction

@@

### Create Plots for Numeric Variables ###
# Create a list to store plots
plots <- list()
# Loop through selected numeric variables and plot their distributions
for (variable in set3_vars) {
  plots[[variable]] <- ggplot(raw_data, aes_string(x = variable)) +
    geom_bar(fill = "#69b3a2", color = "black", alpha = 0.8) +
    labs(title = paste("Distribution of", variable), x = variable, y = "Frequency") +
    theme_minimal(base_size = 30) +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 10)
    )
}
### Combine and Print Plots ###
# Combine all individual plots into one using patchwork with 2 columns and 4 rows
combined_plot <- wrap_plots(plots, ncol = 2, nrow = 2) +
  plot_annotation(
    theme = theme(
      plot.title = element_text(hjust = 0.5, size = 14)
    )
  )
# Print the combined plot
print(combined_plot)
```
\newpage
# References
